---
title: "Malaria Cell Classifier"
subtitle: "Biohack Spring 2026"
author: "Amanda Gale, Aamna Sheikh, Zahera Khatoon"
date: "Spring 2026"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---


```{r, echo=F, message=F}
library(ggplot2)
library(dplyr)
```

```{r, echo=F}
df <- read.csv("clean_features.csv")
df_rf <- read.csv("rf_predictions.csv")
df_svc <- read.csv("svc_predictions.csv")
```

```{r, echo=F}
df_rf <- read.csv("rf_predictions.csv")
df_rf <- df_rf %>% 
  mutate(across(c("Predicted_preTune", "Actual", "Prediction_tuned"), as.factor))

df_svc <- read.csv("svc_predictions.csv")
df_rf <- df_svc %>% 
  mutate(across(c("Predicted_preTune", "Actual", "Prediction_tuned"), as.factor))
```

## Introduction
What is the purpose of this project? Where is the data sourced from?


## Summary of Data
The predictive features used to build this model are...
```{r, echo=F}
head(df)
```


## Exploratory Visualization

```{r, echo=F}
ggplot(df, aes(x=perimeter, y="", fill=label)) + 
  geom_boxplot() +
  theme_minimal() +
  scale_fill_manual(values=c('parasitized'="#AA0000", 'uninfected'="#00AA00")) +
  labs(title="Difference in Cell Perimeter by Disease State", fill="Label", x="Perimeter", y="") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
```

```{r, echo=F}
ggplot(df, aes(x=area, y="", fill=label)) + 
  geom_boxplot() +
  theme_minimal() +
  scale_fill_manual(values=c('parasitized'="#AA0000", 'uninfected'="#00AA00")) +
  labs(title="Difference in Cell Area by Disease State", fill="Label", x="Area", y="") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
```

```{r, echo=F}
ggplot(df, aes(x=eccentricity, y="", fill=label)) + 
  geom_boxplot() +
  theme_minimal() +
  scale_fill_manual(values=c('parasitized'="#AA0000", 'uninfected'="#00AA00")) +
  labs(title="Difference in Cell Eccentricity by Disease State", fill="Label", x="Eccentricity", y="") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
```

```{r, echo=F}
ggplot(df, aes(x=mean_intensity, y="", fill=label)) + 
  geom_boxplot() +
  theme_minimal() +
  scale_fill_manual(values=c('parasitized'="#AA0000", 'uninfected'="#00AA00")) +
  labs(title="Difference in Mean Image Intensity by Disease State", fill="Label", x="Mean Intensity", y="") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
```

```{r, echo=F}
ggplot(df, aes(x=var_intensity, y="", fill=label)) + 
  geom_boxplot() +
  theme_minimal() +
  scale_fill_manual(values=c('parasitized'="#AA0000", 'uninfected'="#00AA00")) +
  labs(title="Difference in Image Intensity Variance by Disease State", fill="Label", x="Intensity Variance", y="") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
```


## Data Shaping

- before feature extraction, images were resized, converted to greyscale, and separated from background
- extraction of rgb color intensity was considered in the model but the colors varied too widely among the 27,000 images in the dataset to be a reliable predictor
- after feature extraction, extreme outliers were removed, data was scaled with the scalar fit to the training data to prevent leakage
- while the data set consisted of masked images, building an additional segmentation mask would have been a promising direction for this project given the notable appearance of intense areas in the cell present in only parasitized images

## Model Parameters


### Random Forest

### SVM

## Conclusion

## References


