---
title: "Malaria Cell Classifier"
author: "Sheikh, Aamna"
date: "Spring 2026"
output: html_document
---


```{r}
df <- read.csv("clean_features.csv")
```

# Summary of Data
```{r}
str(df)
```
```{r}
summary(df)
```

# Class Counts
```{r}
table(df$label)
```

# Basic Stats - by class
```{r}
aggregate(cbind(area, perimeter, eccentricity, mean_intensity) ~ label, data = df, mean)
```

# Feature Comparision
# Area
```{r}
boxplot(area ~ label, data = df, main = "Area by Class", col = c("lightblue", "lightpink"))
```

# Perimeter
```{r}
boxplot(perimeter ~ label, data = df, main = "Perimeter by Class", col = c("lightblue", "lightpink"))
```

# Eccentricity
```{r}
boxplot(eccentricity ~ label, data = df, main = "Eccentricity by Class", col = c("lightblue", "lightpink"))
```

# Mean Intensity
```{r}
boxplot(mean_intensity ~ label, data = df, main = "Mean Intensity by Class", col = c("lightblue", "lightpink"))
```

# Correlation Between Features
```{r}
numeric_df <- df[, c("area", "perimeter", "eccentricity", "mean_intensity")]
cor(numeric_df)
```

###########
# Model Evaluation
```{r}
svc <- read.csv("svc_predictions.csv")
rf <- read.csv("rf_predictions.csv")
```

## Confusion Matrix

```{r}
table(svc$Predicted, svc$Actual)
```{r}
table(rf$Predicted, rf$Actual)
```

## Accuracy (simple)

```{r}
mean(svc$Predicted == svc$Actual)

```{r}
mean(rf$Predicted == rf$Actual)
```

## Compare models
```{r}
mean(svc$Predicted == svc$Actual)
mean(rf$Predicted == rf$Actual)
```

## The average feature values differ slightly between parasitized and uninfected cells, indicating that these features may help distinguish between the two classes.